MS Depression Paper, post replication, started 2022 09 21

### 2022-11-04 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_2022110222_post_replication_playing_w_graphs.Rmd
changed names to MS+Dep/MS-Dep

#extract overall volume and diagnosis
df_lesion_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
  dplyr::select(volume_of_mimosa_lesions,depGroupVar) %>%
  filter(depGroupVar != 0) %>% 
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_violin() + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(violin_lesion_volume) #p=0.067

@@

df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

@@

#effect size dep vs nondep network within nondepressed, eff size = 0.75 (large)
msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "Nondepressed"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#effect size dep vs dep network within depressed, eff size = 0.80, large
msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "Depressed"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)

@@

#label scheme
Network_labels <- c("Depression Network", "Nondepression Network")
Network_labels_wrapped <- c("Depression\nNetwork", "Nondepression\n Network")
Diagnosis_labels <- c("Depressed", "Nondepressed")

@@

df_proportion_network_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
dplyr::select(proportion_volume_lost_per_total_network_size_sum, depGroupVar) %>%
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=depGroupVar, y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  scale_x_discrete(labels=c("1" = "NonDepressed", "2" = "Depressed")) + 
  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion (sum total volume lost/Net Vol) ") +
  ylim(0,1) + 
  scale_fill_manual(values=Diagnosis_colors) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_dep_vs_nondep_scaled_by_network_size) #

@@

bar_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis_summary, aes(x=depGroupVar, y=mean, fill=depGroupVar)) +
  geom_bar(stat = "identity", fill=Diagnosis_colors_faded, color = Diagnosis_colors, size = 1.5, width = 0.8) +  
  geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.2, size=1.5, color = 'black') + #, position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +
  scale_x_discrete(labels=c("Non-depressed" = "Nondepressed", "Depressed" = "Depressed")) + 
#  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion of Whole Brain White Matter Volume Affected") +
  ylim(0,0.6) + 
  theme(plot.title = element_text(size=14, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=14),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(bar_dep_vs_nondep_scaled_by_network_size) #

@@

#geom_bar
interaction_ggplot_bar <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network,  group = interaction(Diagnosis, Network))) + 
    geom_col(position = position_dodge(width=0.6), fill = c(Network_colors_faded, Network_colors_faded), color = c(Network_colors, Network_colors), width = 0.5, size = 1.5) + 
    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.6), width = 0.2,size = 1.5, color='black') +
    scale_fill_manual(values=Network_colors) + #, labels = c("Depression Network", "Nondepression Network"))+ 
    ylim(0,0.6) + 
    xlab("Diagnosis") + 
    ylab("Proportion of White Matter Network Affected") + 
    theme(axis.text.x = element_text(size=14, color="black"),
          axis.text.y = element_text(size=14, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =14), 
          plot.title = element_text(size=16, hjust = 0.5),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 7/5,
          panel.background = element_rect(colour = "white", fill = "white"))+

    geom_signif(
      y_position = c(0.48, 0.48, 0.53, 0.58), 
      xmin = c(0.85, 1.85, 0.85, 1.15),
      xmax = c(1.15, 2.15, 1.85, 2.15),
      annotation = c("***", "***", "*", "NS"), 
      tip_length = 0.03, color = "black",
      size = 1.1, textsize = 5) #+
  #  ggtitle("Diagnosis x Depression Network Disease")
print(interaction_ggplot_bar)

@@

scatter_effsize_val_by_dep_net_overlap_log <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize, color=dep_vs_healthy_effsize)) +
  geom_point(size=2) + 
  scale_color_gradient(low = "#FFFFFF", high = "#ff00ff") +
  geom_smooth(formula = y ~ x, method=lm, color="black")+
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(0, 10, 100, 1000, 10000, 100000)) + 
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (r)") + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=16, color = "black"),
        axis.title = element_text(size=18, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank(),
        legend.position = "none",
        aspect.ratio = 4/4)
print(scatter_effsize_val_by_dep_net_overlap_log)

@@

#### tertile
df_empi_dx_network_proportions_tertiles <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_top_tertile_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_bottom_tertiles_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

@@

#### quintile
df_empi_dx_network_proportions_quintiles <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "Nondepressed", "Depressed")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_top_quintile_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_bottom_quintiles_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

>>>>

#extract overall volume and diagnosis
df_lesion_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
  dplyr::select(volume_of_mimosa_lesions,depGroupVar) %>%
  filter(depGroupVar != 0) %>% 
  mutate(depGroupVar = ifelse(depGroupVar == 1, "MS-Depression", "MS+Depression"))

violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=factor(depGroupVar, level = c('MS+Depression', 'MS-Depression')), y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_violin() + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
 # scale_x_discrete(labels=c("1" = "MS+Depression", "2" = "MS-Depression")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(violin_lesion_volume) #p=0.067

@@

df_empi_dx_network_proportions <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_dep_net_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_nondep_net_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

@@

#effect size dep vs nondep network within nondepressed, eff size = 0.75 (large)
msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "MS-Depression"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)

#effect size dep vs dep network within depressed, eff size = 0.80, large
msdisease_within_versus_outside_depression_network_nondep_group_sums_scaled_by_size_wilcoxon_eff_size <- rstatix::wilcox_effsize(df_empi_dx_network_proportions[(df_empi_dx_network_proportions$Diagnosis == "MS+Depression"),], formula = proportion_affected ~ Network, paired = T, conf.level = 0.95, ci.type = "perc", nboot = 1000)

@@

#label scheme
Network_labels <- c("Depression Network", "Nondepression Network")
Network_labels_wrapped <- c("Depression\nNetwork", "Nondepression\n Network")
#Diagnosis_labels <- c("MS+Depression", "MS-Depression")
Diagnosis_labels <- c("MS+Depression", "MS-Depression")

@@

df_proportion_network_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
dplyr::select(proportion_volume_lost_per_total_network_size_sum, depGroupVar) %>%
   mutate(depGroupVar = ifelse(depGroupVar == 1, "MS-Depression", "MS+Depression"))

 # mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "MS+Depression"))

violin_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis, aes(x=factor(depGroupVar, level = c('MS+Depression', 'MS-Depression')), y=proportion_volume_lost_per_total_network_size_sum, fill=depGroupVar)) +
  geom_violin() +  
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + 
  #scale_x_discrete(labels=c("1" = "MS-Depression", "2" = "MS+Depression")) +

#  scale_x_discrete(labels=c("1" = "MS-Depression", "2" = "MS+Depression")) + 
  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion (sum total volume lost/Net Vol) ") +
  ylim(0,1) + 
  scale_fill_manual(values=Diagnosis_colors) + 
  theme(plot.title = element_text(size=16, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(violin_dep_vs_nondep_scaled_by_network_size) #

@@

#factor(depGroupVar, level = c('MS+Depression', 'MS-Depression'))
bar_dep_vs_nondep_scaled_by_network_size <- ggplot(df_proportion_network_volume_and_diagnosis_summary, aes(x=depGroupVar, y=mean, fill=depGroupVar)) +
  geom_bar(stat = "identity", fill=Diagnosis_colors_faded, color = Diagnosis_colors, size = 1.5, width = 0.8) +  
  geom_errorbar(aes(ymin=mean-se,ymax=mean+se), width=0.2, size=1.5, color = 'black') + #, position=position_dodge(width=0.5), width = 0.2,size = 1.5, color='black') +

  #scale_x_discrete(labels=c("Non-depressed" = "MS-Depression", "MS+Depression" = "MS+Depression")) + 
#  ggtitle(paste0("Proportion Fascicle Volume Affected by Diagnosis")) +
  ylab("Proportion of Whole Brain White Matter Volume Affected") +
  ylim(0,0.6) + 
  theme(plot.title = element_text(size=12, hjust = 0.55),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=14),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))
print(bar_dep_vs_nondep_scaled_by_network_size) #

@@

#geom_bar

#factor(Diagnosis, level = c('MS+Depression', 'MS-Depression'))
interaction_ggplot_bar <- ggplot(data = df_interaction_fit_upper_lower, mapping = aes(x = Diagnosis, y = model_data$predicted, color = Network,  group = interaction(Diagnosis, Network))) + 
    geom_col(position = position_dodge(width=0.6), fill = c(Network_colors_faded, Network_colors_faded), color = c(Network_colors, Network_colors), width = 0.5, size = 1.5) + 
    geom_errorbar(aes(ymin=model_data$conf.low,ymax=model_data$conf.high), position=position_dodge(width=0.6), width = 0.2,size = 1.5, color='black') +
    scale_fill_manual(values=Network_colors) + #, labels = c("MS+Depression", "MS-Depression"))+ 
    ylim(0,0.6) + 
    xlab("Diagnosis") + 
    ylab("Proportion of White Matter Network Affected") + 
    theme(axis.text.x = element_text(size=14, color="black"),
          axis.text.y = element_text(size=14, color="black"),
          axis.title.x = element_blank(), #element_text(size =16), 
          axis.title.y = element_text(size =14), 
          plot.title = element_text(size=16, hjust = 0.5),
          axis.line = element_line(size = 1.5, colour = "black", linetype=1),
          aspect.ratio = 7/5,
          panel.background = element_rect(colour = "white", fill = "white"))+

    geom_signif(
      y_position = c(0.48, 0.48, 0.53, 0.58), 
      xmin = c(0.85, 1.85, 0.85, 1.15),
      xmax = c(1.15, 2.15, 1.85, 2.15),
      annotation = c("***", "***", "*", "NS"), 
      tip_length = 0.03, color = "black",
      size = 1.1, textsize = 5) #+
print(interaction_ggplot_bar)

@@

scatter_effsize_val_by_dep_net_overlap_log <- ggplot(data=df_fascicle_dep_mask_volume_overlap_effsize_value, aes(x=volume_in_dep_mask, y=dep_vs_healthy_effsize, color=dep_vs_healthy_effsize)) +
  geom_point(size=2) + 
  scale_color_gradient(low = "#FFFFFF", high = "#ff00ff") +
  geom_smooth(formula = y ~ x, method=lm, color="black")+
  scale_x_continuous(trans=scales::pseudo_log_trans(base = 10), breaks=c(0, 10, 100, 1000, 10000, 100000)) + 
  xlab("Volume of Fascicle in Depression Network") + 
  ylab("Effect Size (r)") + 
  theme(plot.title=element_text(hjust=0.5),
        axis.text = element_text(size=16, color = "black"),
        axis.title = element_text(size=18, color = "black"),
        axis.line = element_line(color="black", size=1.3),
        panel.background = element_blank(),
        legend.position = "none",
        aspect.ratio = 4/4) + 
  annotate("text", x=1.5, y=0.14, size = 5, 
           label = paste0("P = ", round(summary(dep_overlap_x_dep_vs_healthy_vol_effsize)$coef[2,4],3)))

print(scatter_effsize_val_by_dep_net_overlap_log)

@@

#### tertile
df_empi_dx_network_proportions_tertiles <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_top_tertile_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_bottom_tertiles_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()

@@

#### quintile
df_empi_dx_network_proportions_quintiles <- dep_and_healthy_groups_for_ICD_analysis %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = ifelse(depGroupVar==1, "MS-Depression", "MS+Depression")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
  mutate(Dep_network = proportion_volume_lost_top_quintile_by_network_size_sum) %>%
  mutate(Nondep_network = proportion_volume_lost_bottom_quintiles_by_network_size_sum) %>%
  dplyr::select(subject, Diagnosis, Dep_network, Nondep_network) %>% 
  gather(Network, proportion_affected, -subject, -Diagnosis) %>% #this is like a melt. Keeps empi and depGroupVar out of the melt
  mutate(Network = as.factor(Network)) %>%
  ungroup()
